---
title: "Topic Model Evaluation"
author: "Andreas Blätte (andreas.blaette@uni-due.de)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: topicmodelling.bibtex
vignette: >
  %\VignetteIndexEntry{Topic Model Evaluation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, eval = TRUE}
library(biglda)
library(polmineR)
library(parallel)
library(purrr)
library(stringi)
```


## About

This is a minimal example to demonstrate a workflow for fitting a set of LDA topic
models, and evaluating the models. Note that given the stochastic nature of 
LDA topic models, results are different each time this vignett is built.

We use the REUTERS corpus included in the RcppCWB package here to provide a 
minimal working example.

```{r use_reuters}
use("RcppCWB", corpus = "REUTERS")
```

Given the very limited size of the corpus (`r size(corpus("REUTERS"))` tokens /
`r length(s_attributes("REUTERS", "id"))` documents) , results obtained here are
not robust. 


## Settings

```{r, eval = TRUE}
if (!mallet_is_installed()) mallet_install() # install Mallet if necessary
cores_to_use <- detectCores() - 1L # use all available cores but one
```


## Prepare data

```{r make_instance_list, message = FALSE, eval = TRUE}
discard <- c(tm::stopwords("en"), capitalize(tm::stopwords("en")))

instance_list <- corpus("REUTERS") %>%
  split(s_attribute = "id") %>%
  get_token_stream(p_attribute = "word", subset = {!word %in% discard}) %>%
  lapply(tolower) %>%
  sapply(stri_c, collapse = "\n") %>% # stri_c() is faster than paste0()
  as.instance_list()
```


## Fit models

```{r fit_models, eval = TRUE, message = FALSE, fig.width = 10}
metrics <- lapply(
  seq(from = 6, to = 30, by = 2),
  function(k){
    message("... fitting model for k: ", k)
    BTM <- BigTopicModel(n_topics = k, alpha_sum = 5.1, beta = 0.1)
    BTM$addInstances(instance_list)
    BTM$setNumThreads(cores_to_use)
    BTM$setTopicDisplay(0L, 0L) # no intermediate report on topics
    BTM$logger$setLevel(rJava::J("java.util.logging.Level")$OFF) # remain silent
    BTM$setNumIterations(1000L)
    BTM$estimate()
    lda <- as_LDA(BTM, verbose = FALSE)
    N <- BTM$getDocLengthCounts()
    data.frame(
      k = k,
      cao2009 = BigCao2009(X = B(lda)),
      arun2010 = BigArun2010(beta = B(lda), gamma = G(lda), doclengths = N),
      deveaud2014 = BigDeveaud2014(beta = B(lda))
    )
  }
)
df <- do.call(rbind, metrics)
```


## Plot metrics

This is a minimal version of the much nicer plotting the ldatuning package
offers. Use ldatuning functionality to obtain a ggplot2 plot.

```{r normalize_metrics}
for (m in c("cao2009", "arun2010", "deveaud2014")){
  df[[m]] <- (df[[m]] - min(df[[m]])) / max(df[[m]] - min(df[[m]]))
}
```

```{r metrics_to_minimize, fig.width = 7}
par(mfrow = c(1,2))
plot(
  x = df$k, y = df$cao2009,
  ylab = "normalized metric", xlab = "k (number of topics)",
  main = "Metrics to minimize",
  type = "n"
)
lines(x = df$k, y = df$arun2010, col = "steelblue", lwd = 1)
points(x = df$k, y = df$arun2010, col = "steelblue", pch = 18)
lines(x = df$k, y = df$cao2009, col = "chartreuse3", lwd = 1)
points(x = df$k, y = df$cao2009, col = "chartreuse3", pch = 17)
legend(
  x = "topleft",
  legend = c("arun2010", "cao2009"),
  fill = c("steelblue", "chartreuse3"),
  cex = 0.5
)

plot(
  x = df$k, y = df$cao2009,
  ylab = "normalized metric", xlab = "k (number of topics)", 
  main = "Metrics to maximize",
  type = "n"
)
lines(x = df$k, y = df$deveaud2014, col = "coral2", lwd = 1)
points(x = df$k, y = df$deveaud2014, col = "coral2", pch = 19)
legend(x = "topright", legend = "deveaud2014", fill = "coral2", cex = 0.5)
```
